A summary on Delta Debugging and its uses

### 2019-01-13 Paper/article.tex
Added citations.

@@

\begin{description}
  	\item[``reduce to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a failing configuration, this configuration can be used as a new \cfail\ as it is smaller than the current \cfail. Recurse with a partition granularity of 2.
  	\item[``increase to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass. Recurse with a partition granularity of 2.
  	\item[``increase to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass. Recurse with a partition granularity of $n-1$, as the difference changes only by $\Delta_i$ and thus all coarser partitions would result in unresolved tests again.
  	\item[``reduce to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a failing configuration, this configuration can be used as a new \cfail\, as it is smaller than the current \cfail. Recurse with a partition granularity of $n-1$ (s.a.).
  	\item[``increase granularity''] If none of the above cases apply, i.e. the tests all remain unresolved, then the granularity of the partition of $\Delta$ is increased as in \ddp. 
  	\item[``done''] If $n \ge |\Delta|$ the granularity can not be increased further. In that case every $\Delta_i$ only contains one change. If \cpass\ cannot be increased and \cfail\ not be reduced at that point, the difference between them is 1-minimal and they are returned.\\
\end{description}
The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.\\

Isolation has two technical advantages over simplification:
\begin{enumerate}
	\item It takes less tests to find a 1-minimal failure-inducing difference than to find a minimal failure-inducing subset of changes. %TODO quote paper
	This is because isolation can improve the current configurations if either a test fails or a test passes whereas simplification needs a test to fail to reduce the size of the current subset of changes. Simply put, isolation approaches the solution from both sides simultaneously while simplification only works in one way.
	\item The simplification algorithm \ddp\ relies on monotone configurations, i.e. $test(c)=\cmark \Rightarrow (\forall c' \subseteq c : test(c') = \cmark)$ holds for all $c \in \C$. The isolation algorithm \dddiff\ does not need to assume all configurations are monotone. With some adaptations it can still make use of that property if it is given, though. %TODO quote paper
\end{enumerate}

@@

Consider as an example the above example program \reffail\ that crashes GCC. Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly). Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework.} %TODO quote
compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure. \textit{HOWCOME} yields a \href{fig:failcec}{Cause-Effect-Chain} that includes the comparisons at those locations.

\begin{figure}[h!]
	\fbox{
		\begin{minipage}{\textwidth}
			\begin{enumerate}
			\item Execution reaches \textbf{main}. \\Since the program was invoked as ``ccl -O fail.i'' variable \textbf{$argv[2]$} is now \textbf{``fail.i''}
			\item Execution reaches \textbf{combine\_instructions}. \\ Since $argv[2]$ was ``fail.i'', variable \textbf{$*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$} is now \textbf{$\langle new rtx\_def\rangle$}.
			\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}. \\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}.
			\item Execution ends. \\ Since variable $link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}. The program fails.
			\end{enumerate}
		\end{minipage}
	}
	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail.}
	\label{fig:failcec}
\end{figure}

@@

An important part of software testing today is randomized tests. Given a unit under test (UUT) a randomized test case generator will create random sequences of instructions to the UUT with randomized input data. The test driver will use the generated sequences to test the functionality of the UUT even under implausible circumstances. Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases (see Fuzzing).%TODO quote paper

@@

After generating random test cases it would be useful to minimize the failing test cases so they only contain the instructions that are relevant for inducing the failure. This is where Delta Debugging comes to play: By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case. We define the abstract DD concepts in this context more concretely:

@@

Both \ddp\ and \dddiff\ can be used in this scenario. \ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference. \dddiff\ will likely be more efficient %TODO quote
 but it remains a matter of context which result is more useful for the debugging of the program code.\\

To improve the runtime of the test case minimization static slicing can be initially applied to the test cases. In static slicing Read-Write-Data-Dependencies between the instructions are analyzed and only the instructions that could influence the last instruction remain in the test. This method is a static analysis of the test cases, meaning it does not actually need to execute the tests in order to find instructions that are irrelevant for the failure. This gives static slicing a runtime advantage compared to minimization by Delta Debugging. At the same time static slicing is not as effective as DD minimization. A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization. %TODO quote efficient unit test case minimization

@@

>>>>

\ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, that is they can be applied to the baseline or removed from \td\ without causing an inconsistency. \ddp\ has linear time complexity \cite{Zeller:1999:YMP:318774.318946}.

@@

\begin{description}
  	\item[``reduce to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a failing configuration, this configuration can be used as a new \cfail\ as it is smaller than the current \cfail. Recurse with a partition granularity of 2.
  	\item[``increase to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass. Recurse with a partition granularity of 2.
  	\item[``increase to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass. Recurse with a partition granularity of $n-1$, as the difference changes only by $\Delta_i$ and thus all coarser partitions would result in unresolved tests again.
  	\item[``reduce to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a failing configuration, this configuration can be used as a new \cfail\, as it is smaller than the current \cfail. Recurse with a partition granularity of $n-1$ (s.a.).
  	\item[``increase granularity''] If none of the above cases apply, i.e. the tests all remain unresolved, then the granularity of the partition of $\Delta$ is increased as in \ddp. 
  	\item[``done''] If $n \ge |\Delta|$ the granularity can not be increased further. In that case every $\Delta_i$ only contains one change. If \cpass\ cannot be increased and \cfail\ not be reduced at that point, the difference between them is 1-minimal and they are returned.\\
\end{description}
The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}. It is desribed by Zeller as the successor to \ddp\ and is recommended to generally replace \ddp\ \cite{Zeller:2002:SIF:506201.506206}. \\

Isolation has two technical advantages over simplification:
\begin{enumerate}
	\item It takes less tests to find a 1-minimal failure-inducing difference than to find a minimal failure-inducing subset of changes \cite{Zeller:2002:SIF:506201.506206}.
	This is because isolation can improve the current configurations if either a test fails or a test passes whereas simplification needs a test to fail to reduce the size of the current subset of changes. Simply put, isolation approaches the solution from both sides simultaneously while simplification only works in one way.
	\item The simplification algorithm \ddp\ relies on monotone configurations, i.e. $test(c)=\cmark \Rightarrow (\forall c' \subseteq c : test(c') = \cmark)$ holds for all $c \in \C$. The isolation algorithm \dddiff\ does not need to assume all configurations are monotone. With some adaptations it can still make use of that property if it is given, though \cite{Zeller:2002:SIF:506201.506206}.
\end{enumerate}

@@

Consider as an example the above example program \reffail\ that crashes GCC. Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly). Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework \cite{Zeller:2002:ICC:587051.587053}.}
compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure. \textit{HOWCOME} yields a \href{fig:failcec}{Cause-Effect-Chain} that includes the comparisons at those locations.

\begin{figure}[h!]
	\fbox{
		\begin{minipage}{\textwidth}
			\begin{enumerate}
			\item Execution reaches \textbf{main}. \\Since the program was invoked as ``ccl -O fail.i'' variable \textbf{$argv[2]$} is now \textbf{``fail.i''}
			\item Execution reaches \textbf{combine\_instructions}. \\ Since $argv[2]$ was ``fail.i'', variable \textbf{$*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$} is now \textbf{$\langle new rtx\_def\rangle$}.
			\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}. \\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}.
			\item Execution ends. \\ Since variable $link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}. The program fails.
			\end{enumerate}
		\end{minipage}
	}
	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail \cite{Zeller:2002:ICC:587051.587053}.}
	\label{fig:failcec}
\end{figure}

@@

An important part of software testing today is randomized tests. Given a unit under test (UUT) a randomized test case generator will create random sequences of instructions to the UUT with randomized input data. The test driver will use the generated sequences to test the functionality of the UUT even under implausible circumstances. Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases.

@@

After generating random test cases it would be useful to minimize the failing test cases so they only contain the instructions that are relevant for inducing the failure. This is where Delta Debugging comes to play (\cite{Lei:2005:MRU:1104997.1105255}): By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case. We define the abstract DD concepts in this context more concretely:

@@

Both \ddp\ and \dddiff\ can be used in this scenario. \ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference. \dddiff\ will likely be more efficient (\cite{Zeller:2002:SIF:506201.506206})
 but it remains a matter of context which result is more useful for the debugging of the program code.\\

To improve the runtime of the test case minimization static slicing can be initially applied to the test cases. In static slicing Read-Write-Data-Dependencies between the instructions are analyzed and only the instructions that could influence the last instruction remain in the test. This method is a static analysis of the test cases, meaning it does not actually need to execute the tests in order to find instructions that are irrelevant for the failure. This gives static slicing a runtime advantage compared to minimization by Delta Debugging. At the same time static slicing is not as effective as DD minimization. A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization \cite{Leitner:2007:EUT:1321631.1321698}.

@@

\bibliography{article}
